---
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

:warning: this package is under development - use with care

`qtl2helper` provides with some helper functions to work with objects created by several functions in the [`R/qtl2`](https://kbroman.org/qtl2/) package. 

To install the package do:

```
# make sure the remotes package is installed
remotes::install_github("tavareshugo/qtl2helper")
```

## Quick overview

Let's start with an example dataset from the `qtl2` documentation:

```{r}
library(qtl2)
library(qtl2helper)

# example from R/qtl2 documentation
iron <- read_cross2( system.file("extdata", "iron.zip", package="qtl2") )
map <- insert_pseudomarkers(iron$gmap, step=1)
pr <- calc_genoprob(iron, map, error_prob=0.002)
Xcovar <- get_x_covar(iron)

# QTL scan
out <- scan1(pr, iron$pheno, Xcovar=Xcovar)

# Estimated QTL effects
c2eff <- scan1coef(pr[,"2"], iron$pheno[,"liver"])
```

`qtl2helper` provides methods for the `tidy` function (from the [`broom`]() package), which convert several `R/qtl2` objects to (long) data frames.

```{r}
# convert scan1 object to a tibble
tidy(out)
```

You can provide a `map` of markers to `tidy` to get marker positions in the output:

```{r}
# convert scan1 object to a tibble
tidy(out, map)
```

This makes it ideal to do further plotting, for example with `ggplot2`:

```{r}
library(ggplot2)

out_tbl <- tidy(out, map)

ggplot(out_tbl, aes(x = pos, y = LOD, colour = pheno)) +
  geom_line() +
  facet_grid(. ~ chrom, scales = "free_x", space = "free_x") +
  theme_classic() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(x = "Chromosome", y = "LOD score", colour = "Tissue")
```

Here's the same idea applied to the estimated QTL effects:

```{r}
# tidy the scan1coef object
c2eff_tbl <- tidy(c2eff, map)

# plot
ggplot(c2eff_tbl, aes(x = pos, y = estimate, colour = coef)) +
  geom_line() +
  labs(x = "Chromosome 2 position", y = "QTL effects", colour = "Genotype")
```

